HUME: Human UCCA-based Machine-Translation Evaluation

### 2016-03-09 paper/main.tex
iaa figures
background

While the measure addresses the two major shortcomings
detailed above, it is unable to capture semantic differences which do not show up in
the verb-argument structure, including several very frequent phenomena, 
Another shortcoming of MEANT is its reliance on semantically annotating MT output,
which is often unintelligible. Finally, measures based on a comparison to a

This problem can be addressed by using scores that decompose over the sub-parts of the sentence. 
However, common decompositions into n-grams or even individual words, as employed by the
popular BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005},
are most suitable for evaluating shallow MT approaches and are
less instructive for more structurally-aware methods, which are at
the forefront of MT research today.

These difficulties have motivated the definition of a human measure that decomposes over
the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
this type is the MEANT measure, and its human variant HMEANT
\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
similarity between the output translation and a reference translation in terms of the similarity
between their verbal argument structures
While the measure addresses the two major shortcomings
detailed above, it is unable to capture semantic differences which do not show up in
the verb-argument structure, including several very frequent phenomena, such as
inter-clause linkage, nominalizations and copula clauses
(see \secref{sec:hmeant_comp}).
Another shortcoming of MEANT is its reliance on semantically annotating MT output,
which is often unintelligible. Finally, measures based on a comparison to a
reference translation
are necessarily incomplete, as they compare to a small number of references
out of the the huge number of possible translations a sentence may have
\cite{dreyer2012hyter}.

\paragraph{Machine Translation Evaluation.}
MT evaluation measures has been traditionally pursued through human-based
evaluation that quantifies the adequacy of the translation in preserving the meaning
of the source sentence, and its fluency as a sentence in the target language.
Human evaluation is still accepted as the standard evaluation measure, and is
estimated either in terms of the relative ranking of the outputs of multiple systems
(as is standard in the WMT shared tasks \cite[inter alia]{wmt14}, or by assigning
indiviaul adequacy/fluency scores for each sentence, recently streamlined by \newcite{graham2015}.
The problem with these measures is that they are costly to conduct on large scale, as
they require a human in the loop.

While providing the cleanest signal, human-based evaluation are not ideal for cases where
a wide variety of systems needs to be evaluated (e.g., with many different settings of hyper-parameters),
which has motivated the construction of cheaper methods that correlate with human
judgement. A common approach is to compare the MT output with reference translations,
created or verified by an expert. Methods often rely on some sort of string matching
methods, such as overlap of n-grams, as with the BLEU and METEOR measures, or edit
distance. While these measures greatly supported MT research by allowing to cheply
conduct many MT experiments, they suffer from two principal problems.
First, they rely on comparison to reference translations, which reflect a small
number out of the enormous number of ways to translate a sentence and are thus
inherently incomplete. Second, they decompose over the words and n-grams of the sentence,
and not over their lexical units, and are thus better suited for localizing errors
when evaluating shallow MT methods, and are less telling for identifying the semantic
nature of translation errors in the output.

To address the second issue, more structure-aware methods were proposed, with the
purpose of quantifying how different the reference and translation output are in terms
of their {\it structure}. \newcite{LiuGildea2005} took a syntactic approach, checking
syntactic similarity between the reference and translation\oa{verify!}. \newcite{Owczarzak2008}
took a similar approach using lexical-functional grammar structures.

Comparison of semantic structures with a reference translation
was conducted by \newcite{gimenezMarquez2007,gimenezMarquez2008},
who incorapoted them as features that were combined with lexical and morphosyntactic features.
Perhaps the most notable attempt at a semantic MT evaluation measure is MEANT \cite{lo2011meant} and
its human variant HMEANT, which compares the reference and translation in terms of their semantic
role labeling structure. The measure shares much of HUME's motivation for providing a measure that
decomposes over semantic units, rather than strings. Nevertheless, HMEANT suffers from several major
shortcomings, in that it is not able to capture much frequent phenomena that is not expressed
in the semantic structures, in addition to not addressing the problems incurred by using references.
HUME addresses both of these shortcomings, see \secref{sec:hmeant_comp}.

% semantic representation
\paragraph{Semantic Representation.}
Much interest has been devoted recently to broad-coverage semantic analysis of text.
The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, it is not optimal for our purposes as they do not ground the semantic symbols
in the text, thus not straightwardly allowing a decomposition of the sentence into sub-units,
and are more fine-grained and consequently more difficult to annotate.

Several approaches have been proposed for grounded semantic representations,
generally viewing semantic structures as bi-lexical dependencies. Examples include
the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well as dependencies derived
from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
Despite their appeal as deep and intuitive means for semantic representations,
these approaches also provide fine-grained distinctions and require linguistic expertise to annotate.
%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
%but to our knowledge no results of this brand have been reported.

UCCA (Universal Conceptual Cognitive Annotation) is a natural pick for purposes of MT evaluation
for a number of reasons. 
First, it diverges from the above approaches in representing
semantic structure as a hierarchy of units, defined using coarse-grained and intuitive definitions,
and consequently can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
by building on an established framework for typological description called
``Basic Linguistic Theory'' \cite[BLT]{Dixon:10a,Dixon:10b,Dixon:12},
one of the most widely used frameworks for linguistic description.
Indeed, the scheme's cross-linguistic applicability has been demonstrated through its
application to annotation in English, French, German and Czech. 
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2012}. This is an important feature for the semantic annotation, as it allows us to
assume that the semantic units of the source indeed reflect a common layer of representation that
should be preserved in translation, and so should be represented in the translation output too.
The non-stability of verbal argument structures, which may be translated to a nominal or adjectival
argument structure in another language, is one of the shortcomings of MEANT which HUME addresses.

>>>>

%While the measure addresses the two major shortcomings 
%detailed above, it is unable to capture semantic differences which do not show up in 
%the verb-argument structure, including several very frequent phenomena, 
%Another shortcoming of MEANT is its reliance on semantically annotating MT output,
%which is often unintelligible. Finally, measures based on a comparison to a

%These difficulties have motivated the definition of a human measure that decomposes over
%the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
%this type is the MEANT measure, and its human variant HMEANT
%\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
%similarity between the output translation and a reference translation in terms of the similarity
%between their verbal argument structures
%While the measure addresses the two major shortcomings
%detailed above, it is unable to capture semantic differences which do not show up in
%the verb-argument structure, including several very frequent phenomena, such as
%inter-clause linkage, nominalizations and copula clauses
%(see \secref{sec:hmeant_comp}).
%Another shortcoming of MEANT is its reliance on semantically annotating MT output,
%which is often unintelligible. Finally, measures based on a comparison to a
%reference translation
%are necessarily incomplete, as they compare to a small number of references
%out of the the huge number of possible translations a sentence may have
%\cite{dreyer2012hyter}.

\paragraph{Machine Translation Evaluation.}
MT evaluation measures has been traditionally pursued through human-based
evaluation that quantifies the adequacy of the translation in preserving the meaning 
of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
Human evaluation is still accepted as the standard evaluation measure, and is
estimated either in terms of the relative ranking of the outputs of multiple systems
as is standard in the WMT shared tasks, e.g., \cite{bojar2015findings}, or by assigning
indiviaul adequacy/fluency scores for each sentence, a procedure recently improved
by \newcite{graham2015improving}.

While providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
and is unfeasible for a wide variety of systems or hyper-parameter settings are to be evaluated.
which has motivated the construction of cheaper methods that correlate with human
judgement.

Scalability has been addressed by scores that decompose over the sub-parts of
the translation, and quantify how many of these sub-parts appear in a reference translation,
whose validity has been manually confirmed. The most commonly used measures decompose
over n-grams or even individual words, with examples such as
the BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005}.
Other methods employ a string edit distance for determining the similarity between the
reference and translation \cite{snover2006study}.
While these measures stimulated much progress in MT research by allowing
to evaluate experiments no a massive scale, they 
decompose over the words and n-grams of the sentence,
and are thus better suited for localizing errors
when evaluating shallow MT methods, and are less telling for identifying the semantic
nature of translation errors in the output.

%These methods are most suitable for evaluating shallow MT approaches and are
%less instructive as to what aspects of the reference's meaning were not preserved
%in the translation output.
%A common approach is to compare the MT output with reference translations,
%created or verified by an expert. Methods often rely on some sort of string matching
%methods, such as overlap of n-grams, as with the BLEU and METEOR measures, or edit
%distance.
%First, they rely on comparison to reference translations, which reflect a small
%number out of the enormous number of ways to translate a sentence and are thus
%inherently incomplete.

To address this shortcoming, more structure-aware methods were proposed, with the 
purpose of quantifying how different the reference and translation output are in terms 
of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
quantifying the similarity between the syntactic structures of the reference and translation.
\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
\newcite{gimenez2007linguistic} proposed an evaluation measure that incorapotes features
that capture the overlap between the output and reference in terms of their
semantic (argument structure phenomena), lexical and morphosyntactic features.

Perhaps the most notable attempt at a semantic MT evaluation measure is MEANT and
its human variant HMEANT, which compares the reference and translation in terms of the similarity
of their semantic role labeling structure. HMEANT shares much of HUME's motivation for providing a
measure that decomposes over semantic units, rather than strings, but focuses on verbal
argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
relations and copula clauses. HUME further diverges from both MEANT and the more shallow BLEU and
METEOR measures, in not requiring a reference, as it compares the source directly with the output
translation. We see this difference as crucial where semantic analysis is concerned, as it
is often impossible to assign a semantic structure to a low quality MT output sentence.
See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
and HMEANT.

\paragraph{Semantic Representation.}
UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
whose leaves correspond to the words of the text.
The graph’s nodes, called {\sc units} , are either terminals or several elements jointly
viewed as a single entity according to some semantic or cognitive consideration. Edges bear one
or more categories, indicating the role of the sub-unit in the relation the unit represents. UCCA
currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
action or a state which is persistent in time. Each Scene contains one main relation, as well
as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
contains two Scenes, whose main relations are ``graduating'' and ``John''. The participant ``John''
is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
relations between Scenes and the internal structures of complex participants and relations.\oa{add figure}

% UCCA Motivation
UCCA is a natural candidate for defining a semantic
MT evaluation measure for a number of reasons. 
First, it defines a coarse-grained and intuitive set of distinctions definitions,
and consequently can be reliably annotated by non-experts, after as little as two hours
of training \cite{marinotti2014}.
Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
application to annotation in English, French, German and Czech. 
Third, the scheme has been shown to be semantically stable
in translation: UCCA annotations of translated text usually contain the same set of relationships
\cite{sulem2012}. This is an indication that the scheme reflect a common layer of representation that
should be preserved in translation, and so should be represented in the translation output too.
We note that that verbal argument structures are less stable in this sense, as a verbal
argument structure may be translated or paraphrased to a nominal or adjectival
argument structure. For example, ``after he graduated, Tom moved to NYC'' conveys a similar meaning
to ``after graduation, Tom moved to NYC''.

While most broad-coverage semantic work has focused on shallow semantic structures,
predominantly argument structures annotated with their semantic roles, much recent
interest has been devoted to defining more complete semantic representation schemes.
The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
shares much of UCCA's motivation for defining a more complete semantic annotation.
However, using AMR is not optimal for defining a decomposition of a sentence into semantic
units as it does not ground its semantic symbols in the text,
thus not straightwardly allowing a decomposition of the sentence into sub-units.
Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.

Other approaches have been proposed for semantic representations, which are grounded in the words
and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
Examples include the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well
as dependencies derived from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
Nevertheless, despite their appeal as deep and intuitive means for semantic representations,
these approaches require linguistic expertise for their annotatation, due to the intricate and
fine-grained distinctions they provide, and are thus in our view less suitable and
easily interpretable means for our purposes.
%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
%but to our knowledge no results of this brand have been reported.
